The Country-level Impact of Educational Patterns on the Rate of Entrepreneurship

### 2014-10-24 EducProposal.Rmd
Minor edits to proposal and revised README

The process of allocating public funds for education is an opportunity to make calculated long-term investments in a nation's populace.  The rationale for decisions made during this process, however, is often difficult to understand.  For example, while many claims have been made in entrepreneurship literature regarding the importance of cultural and educational factors -- not just structural and economic factors -- in determining individual and societal tendencies to start a business, these factors are often not included in the decisionmaking process. As a matter of fact, only a few empirical studies have actually been carried out on this topic and most of them focused on which attitudes were favourable to starting a business, while almost none explored how cultural patterns affected these dispositions.

@@

Drawing a comparison between France and Britain, to what extent do the behavioural characteristics of the education system - individualistic, group-oriented, and more or less positive reinforcement in teaching - affect the rate of entrepreneurship in each country?

@@

Over a hundred years have passed since Max Weber tried to explain how religious-ethical motivations affect entrepreneurship [@weberprotestant], and religion may be less effective in justifying why some are more likely to start businesses than others. Still, that was a founding stone and it is now widely admitted that “cultural features influence attitudes towards start-ups and that these attitudes, in turn, have an impact on start-up activities” [@OECD2009].

Despite various claims of the existence of a cultural effect, few empirical studies have actually been carried out. It took a few decades more before Davidsson, along with Delmar at first (1992), then alone (1995), then with Wiklund [@davidsson1997values], at his turn, laid a stepping stone by investigating the relationship between structural factors, mainly economic, and cultural factors. As they compared two Swedish regions similar in terms of structure, and observed different outcomes, they concluded on the possibility of a small cultural effect.

This thesis is an attempt at finding out if different patterns in education lead to different outcomes in terms of start-up activities. Indeed, like many in the education literature tend to assert, we believe there is an effect by school personnel on student self-esteem [@Scott1996]. We will explore further the way this mechanism works. As for entrepreneurship literature, it generally acknowledges the influence of self-confidence on start-up activities although “self-confidence in one’s own capabilities to start a business has been found to depend almost exclusively on the individual features of the respective person and his or her integration into social networks” [@OECD2009]. We intend to show that these individual characteristics are partly shaped at an institutional and collective level through the education systems and staff attitudes.

Of course, King and Sobel [@Kerry2008], among others, have already tried to show how school choice can increase the rate of youth entrepreneurship, but their perspective was slightly different. They  asserted that a more business-like environment in K-12 education was the key, which was favoured by voucher programmes. We believe there is a more widely spread systematic effect of cultural and educational patterns. David Watkins [@Watkins2010] cross-cultural perspective on teaching and learning shall be useful in that respect. Kai-Ming Cheng and Kam-Cheung Wong’s study about school effectiveness in East Asia [@Cheng2006] could also be of help.

Also, as self-confidence is considered to rely partly on one’s integration within social networks [@OECD2009], we will refer to cross-cultural studies such as that of Green, Deschamps and Páez [@green2005variation] about individualism and collectivism, what Freeman [@freeman1997demographic] called idiocentrism and allocentrism. Raeff, Greenfield and Quiroz [@raeff2000conceptualizing] have also conceptualised interpersonal relationships in both contexts. More school-focused, Yamauchi [@yamauchi1998individualism] as well as Davies and Aurini [@davies2003homeschooling] have looked at individualism and collectivism in the classroom and in pedagogy respectively. All these will be important on the way.

@@

Stronger culture of positive reinforcement ---> higher degree of individual self-esteem ---> higher probability an individual will become an entrepreneur ---> higher rate of entrepreneurship at a country-level

@@

GEM measure of entrepreneurship is often used or refered to in OECD studies.

@@

From PISA 2012, we will use the items regarding engagement, drive and self-beliefs (Table III.2.3.a). More specifically we will compute data regarding the sense of belonging to measure the strength of inclusion in social networks. We will also process measures of self-perception. They all come as a percentage of the country-representative student population.

From World Value Survey, we will use the last data available, that of the fifth wave (2005-2009). Our main issue is that data was collected at different points in time in Britain (2005) and in France (2006), but in both cases before the economic crisis, which might have affected people’s prospects and values differently. We will specifically look at the following items:

V86   How important are adventure and taking risks to the person surveyed?

V119 How good is competition?

@@

To compute these data together is quite a challenge and we are not yet fully sure on how to proceed. But at least, these multiple reliable measures from well-recognised sources should give us a composite but also comprehensive indicator about the average degree of self-confidence of individuals in each country.

@@

As we mentioned that self-confidence also depends on the integration in social networks, we will partly rely on the sense of belonging as defined by OECD PISA study but again we will need to compute an index. In order to do so, we will also rely on the following items from the World Value Survey 2005-2009 wave about “sense of belonging to a group“ and “sense of individuality”.

@@

1. Finance  
2. Government policies  
3. Government programmes  
4. Entrepreneurial Education and Training  
5. Research and Development Transfer (R & D)  
6. Commercial and professional infrastructure  
7. Internal Market openness  
8. Physical infrastructure and services  
9. Cultural and social norms

We are not yet fully sure on how to control for all these though.

@@

Then, should indeed different patterns be observed, we will need to proceed to an Ordinary Least Square (OLS) Regression in order to check to what extent these characteristics affect the rate of entrepreneurship in the country.

@@

1. Self-confidence: It is difficult to obtain a credible measurement of self confidence, but we will compute an index to include some quantiative approach in measuring this factor.

2. Positive reinforcement: Again even if this is not easy to characterize quantify, we will attempt to compute an index and characterize it on a scale. Should we eventually consider there is a lot of positive reinforcement in one country and very little in another, perhaps we could go as far as using a dummy variable.

@@

5. Return on Investment: Future benefits of a euro spent at present. Here return on investment would only start to be visible after some time, once most of a generation has got into active life which may take up to five to ten years. But even then, not all the expected propensity to start a business will be palpable. When you add it to the mechanism of the social discount rate, that huge distance in time as of now seem to prevent any calculations of return on investment when it comes to public budgeting for education policies.

Chart about factors that affect entrepreneurship.
Strengthening entrepreneurship and economic development in East Germany: Lessons from local approaches.

@@

>>>>

The process of allocating public funds for education is an opportunity to make calculated long-term investments in a nation's populace.  The rationale for decisions made during this process, however, is often difficult to understand.  For example, while many claims have been made in entrepreneurship literature regarding the importance of cultural and educational factors -- not just structural and economic factors -- in determining individual and societal tendencies to start a business, these factors are often not included in the decisionmaking process. As a matter of fact, only a few empirical studies have actually been carried out on this topic and most of them focused on which attitudes were favorable to starting a business, while almost none explored how cultural patterns affected these dispositions.

@@

Drawing a comparison between France and Britain, to what extent do the behavioral characteristics of the education system - individualistic, group-oriented, and more or less positive reinforcement in teaching - affect the rate of entrepreneurship in each country?

@@

Over a hundred years has passed since Max Weber tried to explain how religious-ethical motivations affect entrepreneurship [@weberprotestant], and religion may be less effective in justifying why some are more likely to start businesses than others. Still, that was a founding stone and it is now widely admitted that “cultural features influence attitudes towards start-ups and that these attitudes, in turn, have an impact on start-up activities” [@OECD2009].

Despite various claims of the existence of a cultural effect, few empirical studies have actually been carried out. It took a few decades more before Davidsson, along with Delmar at first (1992), then alone (1995), and then with Wiklund [@davidsson1997values], at his turn, laid a stepping stone by investigating the relationship between structural factors -- mainly economic, and cultural factors. As they compared two Swedish regions similar in terms of structure and observed different outcomes, they concluded there may be a small cultural effect in attitudes toward start-up activities.

This thesis is an attempt at finding out whether different patterns in education lead to different outcomes in terms of start-up activities. Indeed, like many in the education literature tend to assert, we believe school personnel have an affect on student self-esteem [@Scott1996]. We will further explore how this mechanism works. As for entrepreneurship literature, it generally acknowledges the influence of self-confidence on start-up activities, although “self-confidence in one’s own capabilities to start a business has been found to depend almost exclusively on the individual features of the respective person and his or her integration into social networks” [@OECD2009]. We intend to show that these individual characteristics are partly shaped at an institutional and collective level through the education systems and staff attitudes.

Of course, King and Sobel [@Kerry2008], among others, have already tried to show how school choice can increase the rate of youth entrepreneurship, but their perspective was slightly different. They  asserted that a more business-like environment in K-12 education was the key, which was favored by voucher programs. We believe there is a more widely spread systematic effect of cultural and educational patterns. David Watkins [@Watkins2010] cross-cultural perspective on teaching and learning shall be useful in that respect. Kai-Ming Cheng and Kam-Cheung Wong’s study about school effectiveness in East Asia [@Cheng2006] could also be of help.

Also, as self-confidence is considered to rely partly on one’s integration within social networks [@OECD2009], we will refer to cross-cultural studies such as that of Green, Deschamps and Páez [@green2005variation] about individualism and collectivism, what Freeman [@freeman1997demographic] called idiocentrism and allocentrism. Raeff, Greenfield and Quiroz [@raeff2000conceptualizing] have also conceptualized interpersonal relationships in both contexts. More school-focused, Yamauchi [@yamauchi1998individualism] as well as Davies and Aurini [@davies2003homeschooling] have looked at individualism and collectivism in the classroom and in pedagogy respectively. All these will be important on the way.

@@

Stronger culture of positive reinforcement ---> higher degree of individual self-esteem ---> higher probability an individual will become an entrepreneur ---> higher rate of entrepreneurship at a country-level.

@@

Note: The GEM measure of entrepreneurship is often used or referred to in OECD studies.

@@

From PISA 2012, we will use the items regarding engagement, drive and self-beliefs (Table III.2.3.a). This data is available in Excel format from the OECD. More specifically, we will compute data regarding the sense of belonging to measure the strength of inclusion in social networks. We will also process measures of self-perception. They all come as a percentage of the country-representative student population.

From World Value Survey, we will use the latest data available, that of the fifth wave (2005-2009). Our main issue is that data was collected at different points in time in Britain (2005) and in France (2006), but in both cases before the economic crisis, which might have affected people’s prospects and values differently. We will specifically look at the following items:

V86   How important are adventure and taking risks to the person surveyed?  
V119 How good is competition?

@@

To compute these data together is quite a challenge and we are not yet fully sure on how to proceed. But at least, these multiple reliable measures from well-recognized sources should give us a composite but also comprehensive indicator about the average degree of self-confidence of individuals in each country.

@@

As we mentioned that self-confidence also depends on the integration in social networks, we will partly rely on the sense of belonging as defined by OECD PISA study, but again we will need to compute an index. In order to do so, we will also rely on the following items from the World Value Survey 2005-2009 wave about “sense of belonging to a group“ and “sense of individuality”.

@@

1. Finance  
2. Government policies  
3. Government programs  
4. Entrepreneurial Education and Training  
5. Research and Development Transfer (R & D)  
6. Commercial and professional infrastructure  
7. Internal Market openness  
8. Physical infrastructure and services  
9. Cultural and social norms

We are not yet fully sure on how to best control for all these variables though and will work on this in the coming weeks.

@@

Then, should indeed different patterns be observed, we will proceed to an Ordinary Least Square (OLS) Regression to determine the extent of the affect of these characteristics on the rate of entrepreneurship in each country.

@@

1. Self-confidence: It is difficult to obtain a credible measurement of self confidence, but we will compute an index to include a quantitative approach in measuring this factor.

2. Positive reinforcement: Again, even if this is not easy to characterize quantitatively, we will attempt to compute an index and characterize it on a scale. Should we eventually consider there is considerable positive reinforcement in one country and very little in another, perhaps we could go as far as using a dummy variable.

@@

5. Return on Investment: Future benefits of a euro spent at present. Here, return on investment would only start to be visible after some time, once most of a generation has got into active life which may take up to five to ten years. But even then, not all the expected propensity to start a business will be palpable. When you add it to the mechanism of the social discount rate, that huge distance in time as of now seem to prevent any calculations of return on investment when it comes to public budgeting for education policies.

6. Individual vs. societal factors: The below chart identifies several of the individual and societal factors impacting the decision to become an entrepreneur.

Strengthening entrepreneurship and economic development in East Germany: Lessons from local approaches.

@@

### Conclusion  
The impact of cultural factors in education on the rate of entrepreneurship in a given country is an important and under-researched topic that should be considered by policymakers constructing public education budgets. Through this project, we hope to encourage the consideration of these cultural factors by providing a qualitative and quantitative method for understanding the impact of these factors. Although several challenges remain to compiling the indexes discussed in this report, we are confident we can produce a helpful framework for examining and understanding these factors.
